rgbif: R client for working with GBIF species occurrence data

### 2015-11-25 manuscript.Rmd
manuscript improvements on occurrences search section, conclusion fleshed out

Perhaps the most fundamental element in many fields of ecology is
the inividual. How many individuals of each species in a given
location forms the basis for many sub-fields of ecology and evolution.
Some research questions necessitate collecting new data, while
others can easily take advantage of existing data. In fact, some
ecology fields are built largely on existing data, e.g., macro-ecology
[@Brown_1995; @Beck_2012].

Data on individuals, including which species, and where they're found,
can be used for a large number of research questions. Biodiversity
records have been used for a suite of other use cases:
validating habitat suitability models with real occurrence data
[@Ficetola_2014]; ancestral range reconstruction [@Ferretti_2015;
@Mendoza_2015]; development of invasive species watch lists
[@Faulkner_2014]; evaluate risk of invasive species spread
[@Di_Febbraro_2013]; and effects of climate change on future
biodiversity [@Brown_2015].

In addition to wide utility, this data is important for conservation.
Biodiversity loss is one of the greatest challenges of our time
[@Pimm_2014]. Some have called this the sixth great mass
extinction [@Ceballos_2015]. Given this challenge there is a great
need for data on specimen records, whether collected from live sightings
in the field or specimens in museums.

There are many online services that collect and maintain specimen records.
However, Global Biodiversity Information Facility (hereafter, GBIF,
http://www.gbif.org/) is the largest collection of biodiversity
records globally, currently with 580 million records, 1.6 million
taxa, 15,000 datasets from 770 publishers (figures collected on 2015-10-04).
Many large biodiversity warehouses such as iNaturalist
(http://www.inaturalist.org/), VertNet (http://vertnet.org/), and USGS's
Biodiversity Information Serving Our Nation (BISON; http://bison.usgs.ornl.gov/)
all feed into GBIF.

Herein, we describe the `rgbif` library [@rgbif] for working with
GBIF data in the R programming environment [@R]. R is a widely used
language in academia, and in non-profit and private sectors.
Importantly, R makes it easy to do all of the steps of the
research process, including data management, data manipulation
and cleaning, statistics, and vizualization. Thus, an R client
for getting GBIF data is a powerful tool to facilitate
reproducible research.

@@

The `rgbif` package is completely written in R, uses an [MIT license][mit]
to maximize use everywhere. `rgbif` is developed publicly on GitHub
at [https://github.com/ropensci/rgbif][ghrgbif], where development versions
of the package can be installed, and bugs and feature requests reported.
Stable versions of `rgbif` can be installed from [CRAN][cranrgbif], the
distribution network for R packages. `rgbif` is part of the rOpenSci project,
a developer network making R software to facilitate reproducible research.

@@

`rgbif` is designed following the [GBIF Application Programming Interface][gbifdocs],
or API. The GBIF API has four major components: registry, taxonomic names,
occurrences, and maps. We ignore maps in `rgbif` as it is concerned with
generating maps primarily for web applications. `rgbif` has a suite of functions dealing
with each of registry, taxonomic names, and occurrences - we'll go through
each in turn describing design and example usage.

@@

With each request `rgbif` makes to GBIF's API, we send request headers that
tell GBIF that the request is coming from `rgbif`, including what version of
the package. This will help GBIF know what proportion of requests are coming from
this package, and therefore R, as most requests likely will come from `rgbif`.

@@

The GBIF reqistry API services are spread across four sets
of functions:

@@

Datasets are owned by organizations. Organizations are endorsed by nodes to share datasets with GBIF.
Datasets are published through institutions, which may be hosted at another organization. A network
is a group of datasets (managed by GBIF).

@@

Also, check out `datasets()` and `dataset_suggest()` for searching for datasets.

@@

Dataset metrics are another useful way of digging in to figure out what
datasets you may want to use. One drawback is that these data are only
available for datasets of type _checklist_, but there are quite a lot of
them (`r dataset_search(type = "checklist")$meta$count`).

@@

The goal of these name functions is often to settle on a taxonomic name
known to GBIF's database. This serves two purposes: 1) when referring to
a taxonomic name, you can point to a URI on the internet, and 2) you can
search for metadata on that species, and occurrences of that species in
GBIF.

Taxonomic names are particularly tricky. Many different organizations have
their own unique codes for the same taxonomic names, and some taxonomic
groups have preferred sources for the definitive names for that group.

@@

One of the quickest ways to search for names is using `name_suggest()`, which
does a very quick search and returns minimal data. Here, we're searching for
the query tem _Pum_, and we get back a bunch of names:

@@

@@

@@

The download API syntax is similar to the occurrence search API
in that the same parameters are used, but the way in which the query
is defined is different. For example, in the download API you can
do greater than searches (i.e., `latitude > 50`), whereas you can not
do that in the occurrence search API. Thus, we can't make the query
interace exactly the same for both search and download functions.

Using the download service can be as few as three steps: 1) Request data
via a search; 2) Download data; 3) Import data into R.

@@

@@

...

@@

```{r}
out <- occ_search(issue='DEPTH_UNLIKELY', limit = 500)
NROW(out)
out %>% occ_issues(-cudc) %>% .$data %>% NROW
```

@@

* pt 1
* pt 2
* pt 3
* pt 4

>>>>

Perhaps the most fundamental element in many fields of ecology is
the individual. The number of individuals of each species in a given
location forms the basis for many sub-fields of ecology and evolution.
Some research questions necessitate collecting new data, while
others can easily take advantage of existing data. In fact, some
ecology fields are built largely on existing data, e.g., macro-ecology
[@Brown_1995; @Beck_2012].

Data on individuals, including which species, and where they're found,
can be used for a large number of research questions. Biodiversity
records have been used for a suite of other use cases:
validating habitat suitability models with real occurrence data
[@Ficetola_2014]; ancestral range reconstruction [@Ferretti_2015;
@Mendoza_2015]; development of invasive species watch lists
[@Faulkner_2014]; evaluating risk of invasive species spread
[@Di_Febbraro_2013]; and effects of climate change on future
biodiversity [@Brown_2015].

In addition to wide utility, this data is important for conservation.
Biodiversity loss is one of the greatest challenges of our time
[@Pimm_2014], and some have called this the sixth great mass
extinction [@Ceballos_2015]. Given this challenge there is a great
need for data on specimen records, whether collected from live sightings
in the field or specimens in museums.

There are many online services that collect and maintain specimen records.
However, Global Biodiversity Information Facility (hereafter, GBIF,
http://www.gbif.org) is the largest collection of biodiversity
records globally, currently with 640 million records, 1.6 million
taxa, 15,000 datasets from 780 publishers (current as of 2015-11-25).
Many large biodiversity warehouses such as iNaturalist
(http://www.inaturalist.org), VertNet (http://vertnet.org), and USGS's
Biodiversity Information Serving Our Nation (BISON; http://bison.usgs.ornl.gov)
all feed into GBIF.

Herein, we describe the `rgbif` software library [@rgbif] for working with
GBIF data in the R programming environment [@R]. R is a widely used
language in academia, and in non-profit and private sectors.
Importantly, R makes it easy to do all of the steps of the
research process, including data management, data manipulation
and cleaning, statistics, and vizualization. Thus, an R client
for getting GBIF data is a powerful tool to facilitate
reproducible research.

@@

The `rgbif` package is nearly completely written in R (a small Javascript
library is included for reading well known text), uses an [MIT license][mit]
to maximize use everywhere. `rgbif` is developed publicly on GitHub
at [https://github.com/ropensci/rgbif][ghrgbif], where development versions
of the package can be installed, and bugs and feature requests reported.
Stable versions of `rgbif` can be installed from [CRAN][cranrgbif], the
distribution network for R packages. `rgbif` is part of the rOpenSci project
(http://ropensci.org), a developer network making R software to facilitate
reproducible research.

@@

`rgbif` is designed following the [GBIF Application Programming Interface][gbifdocs],
or API. The GBIF API has four major components: registry, taxonomic names,
occurrences, and maps. We also include functions to interface with the OAI-PMH
GBIF service; only dataset information is available vis this service, however. We
ignore maps in `rgbif` as it is concerned with generating maps primarily for web
applications. `rgbif` has a suite of functions dealing with each of registry, taxonomic
names, and occurrences - we'll go through each in turn describing design and example usage.

@@

With each request `rgbif` makes to GBIF's API, we send request headers that
tell GBIF that the request is coming from `rgbif`, including what version of
the package. This helps GBIF know what proportion of requests are coming from
this package, and therefore R, as most requests likely will come from `rgbif`;
this information is helpful for them in thinking about how people are using
GBIF data.

@@

The GBIF reqistry API services are spread across five sets of functions
via the main API:

@@

And dataset information in general is available via the OAI-PMH service,
functions in `rgbif` prefixed with `gbif_oai_`.

Datasets are owned by organizations. Organizations are endorsed by nodes to share
datasets with GBIF. Datasets are published through institutions, which may be
hosted at another organization. A network is a group of datasets (managed by GBIF).
Datasets are the units that matter the most with respect to registry information,
while installations, networks, nodes, and organizations are simply higher level
organizational structure.

@@

See also `datasets()` and `dataset_suggest()` for searching for datasets.

@@

Dataset metrics are another useful way of figuring out what
datasets you may want to use. One drawback is that these metrics data are only
available for datasets of type _checklist_, but there are quite a lot of
them (`r dataset_search(type = "checklist")$meta$count`).

@@

The goal of these name functions is often to settle on a taxonomic name
known to GBIF's database. This serves two purposes: 1) when referring to
a taxonomic name, you can point to a URI on the internet, and 2) you can
search for metadata on a taxon, and occurrences of that taxon in
GBIF.

Taxonomic names are particularly tricky. Many different organizations have
their own unique codes for the same taxonomic names, and some taxonomic
groups have preferred sources for the definitive names for that group. That's
why it's best to determine what name GBIF uses, and its associated identifier,
for the taxon of interest instead of simply searching for occurrences with a
taxonomic name.

@@

One of the quickest ways to search for names is using `name_suggest()`, which
does a very quick search and returns minimal data. Here, we're searching for
the query tem _Pum_, and we get back many names:

@@

With these results, you can then proceed to search for occurrences with the
taxon key(s), or drill down further with other name searching functions to
get the exact taxon of interest.

@@

The download service, like the occurrence search service, is rate-limited.
That is, you can only have one to three downloads running simultaneously for
your user credentials. However, simply check when a download job is complete,
then you should be able to start a new download request.

@@

The download API syntax is similar to the occurrence search API
in that the same parameters are used, but the way in which the query
is defined is different. For example, in the download API you can
do greater than searches (i.e., `latitude > 50`), whereas you can not
do that in the occurrence search API. Thus, unfortnately, we couldn't
make the query interace exactly the same for both search and
download functions.

Using the download service can consist of as few as three steps: 1) Request
data via a search; 2) Download data; 3) Import data into R.

@@

#### Search for occurrences

@@

The output of `occ_search()` borrows the tidy `data.frame` idea from the `dplyr`
R package, so that no matter how large the `data.frame`, the output is easily
assessed because only a few of the records (rows) are shown, only a few columns
are shown (with others shown in name only), and metadata is shown on top of
the `data.frame` to indicate data found and returned, media
records found, unique taxonomic hierarchies returned, and the query executed.

The output of these examples, except one, aren't shown, but all run correctly.

Search by species name, using `name_backbone()` first to get key

```{r}
(key <- name_suggest(q = 'Helianthus annuus', rank = 'species')$key[1])
occ_search(taxonKey = key, limit = 2)
```

Instead of getting a taxon key first, you can search for a name directly

```{r eval=FALSE}
occ_search(scientificName = 'Ursus americanus')
```

Search for many species

```{r eval=FALSE}
splist <- c('Cyanocitta stelleri', 'Junco hyemalis', 'Aix sponsa')
keys <- sapply(splist, function(x) name_suggest(x)$key[1], USE.NAMES = FALSE)
occ_search(taxonKey = keys, limit = 5, return = 'data')
```

Spatial search, based on well known text format, or a bounding box set of
four coordinates

```{r eval=FALSE}
# well known text
occ_search(geometry = 'POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))', limit = 20)
# bounding box
occ_search(geometry = c(-125.0,38.4,-121.8,40.9), limit = 20)
```

Get only occurrences with lat/long data

```{r eval=FALSE}
occ_search(hasCoordinate = TRUE, limit = 5)
```

Get only those occurrences with spatial issues

```{r eval=FALSE}
occ_search(hasGeospatialIssue = TRUE, limit = 5)
```

@@

```{r}
out <- occ_search(issue = 'DEPTH_UNLIKELY', limit = 500)
NROW(out)
out %>% occ_issues(-cudc) %>% .$data %>% NROW
```

@@

The `rgbif` R package provides a programmatic interface to GBIF's
application programming interface (API) - a powerful tool for making
research using species occurrence data reproducible. In fact,
the `rgbif` package has already been used in 22 scholarly publications
(as of 2015-11-14).

The `rgbif` package is relatively stable now, and should not have
many breaking changes unless necessitated due to changes in the
GBIF API.

One area of focus in the future is to attempt to solve many use
cases that have been brought up with respect to GBIF data. For
example, some specimens are included in GBIF that are located
in botanical gardens. For many research questions, researchers are
interested in "wild" type occurrences, not those in human curated
scenarios. Making removal of these occurrences easy would be
very useful, but is actually quite a hard problem.

### 2015-11-25 manuscript.tex
manuscript improvements on occurrences search section, conclusion fleshed out

Perhaps the most fundamental element in many fields of ecology is the
inividual. How many individuals of each species in a given location
forms the basis for many sub-fields of ecology and evolution. Some
research questions necessitate collecting new data, while others can
easily take advantage of existing data. In fact, some ecology fields are
built largely on existing data, e.g., macro-ecology (Brown, 1995; Beck
et al., 2012).

Data on individuals, including which species, and where they're found,
can be used for a large number of research questions. Biodiversity
records have been used for a suite of other use cases: validating
habitat suitability models with real occurrence data (Ficetola et al.,
2014); ancestral range reconstruction (Ferretti et al., 2015; Mar{í}a
Mendoza et al., 2015); development of invasive species watch lists
(Faulkner et al., 2014); evaluate risk of invasive species spread
(Febbraro et al., 2013); and effects of climate change on future
biodiversity (Brown et al., 2015).

In addition to wide utility, this data is important for conservation.
Biodiversity loss is one of the greatest challenges of our time (Pimm et
al., 2014). Some have called this the sixth great mass extinction
(Ceballos et al., 2015). Given this challenge there is a great need for
data on specimen records, whether collected from live sightings in the
field or specimens in museums.

There are many online services that collect and maintain specimen
records. However, Global Biodiversity Information Facility (hereafter,
GBIF, \url{http://www.gbif.org/}) is the largest collection of
biodiversity records globally, currently with 580 million records, 1.6
million taxa, 15,000 datasets from 770 publishers (figures collected on
2015-10-04). Many large biodiversity warehouses such as iNaturalist
(\url{http://www.inaturalist.org/}), VertNet
(\url{http://vertnet.org/}), and USGS's Biodiversity Information Serving
Our Nation (BISON; \url{http://bison.usgs.ornl.gov/}) all feed into
GBIF.

Herein, we describe the \texttt{rgbif} library (Chamberlain et al.) for
working with GBIF data in the R programming environment (R Core Team,
2014). R is a widely used language in academia, and in non-profit and
private sectors. Importantly, R makes it easy to do all of the steps of
the research process, including data management, data manipulation and
cleaning, statistics, and vizualization. Thus, an R client for getting
GBIF data is a powerful tool to facilitate reproducible research.

@@

The \texttt{rgbif} package is completely written in R, uses an
\href{http://choosealicense.com/licenses/mit/}{MIT license} to maximize
use everywhere. \texttt{rgbif} is developed publicly on GitHub at
\href{https://github.com/ropensci/rgbif}{\url{https://github.com/ropensci/rgbif}},
where development versions of the package can be installed, and bugs and
feature requests reported. Stable versions of \texttt{rgbif} can be
installed from
\href{https://cran.rstudio.com/web/packages/rgbif/}{CRAN}, the
distribution network for R packages. \texttt{rgbif} is part of the
rOpenSci project, a developer network making R software to facilitate
reproducible research.

@@

\texttt{rgbif} is designed following the
\href{http://www.gbif.org/developer/summary}{GBIF Application
Programming Interface}, or API. The GBIF API has four major components:
registry, taxonomic names, occurrences, and maps. We ignore maps in
\texttt{rgbif} as it is concerned with generating maps primarily for web
applications. \texttt{rgbif} has a suite of functions dealing with each
of registry, taxonomic names, and occurrences - we'll go through each in
turn describing design and example usage.

@@

With each request \texttt{rgbif} makes to GBIF's API, we send request
headers that tell GBIF that the request is coming from \texttt{rgbif},
including what version of the package. This will help GBIF know what
proportion of requests are coming from this package, and therefore R, as
most requests likely will come from \texttt{rgbif}.

@@

Datasets are owned by organizations. Organizations are endorsed by nodes
to share datasets with GBIF. Datasets are published through
institutions, which may be hosted at another organization. A network is
a group of datasets (managed by GBIF).

@@

Also, check out \texttt{datasets()} and \texttt{dataset\_suggest()} for
searching for datasets.

@@

Dataset metrics are another useful way of digging in to figure out what
datasets you may want to use. One drawback is that these data are only
available for datasets of type \emph{checklist}, but there are quite a
lot of them (2483).

@@

One of the quickest ways to search for names is using
\texttt{name\_suggest()}, which does a very quick search and returns
minimal data. Here, we're searching for the query tem \emph{Pum}, and we
get back a bunch of names:

@@

The download API syntax is similar to the occurrence search API in that
the same parameters are used, but the way in which the query is defined
is different. For example, in the download API you can do greater than
searches (i.e., \texttt{latitude \textgreater{} 50}), whereas you can
not do that in the occurrence search API. Thus, we can't make the query
interace exactly the same for both search and download functions.

Using the download service can be as few as three steps: 1) Request data
via a search; 2) Download data; 3) Import data into R.

@@

@@

\ldots{}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{out <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{issue=}\StringTok{'DEPTH_UNLIKELY'}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{500}\NormalTok{)}
\KeywordTok{NROW}\NormalTok{(out)}
\CommentTok{#> [1] 4}
\NormalTok{out %>%}\StringTok{ }\KeywordTok{occ_issues}\NormalTok{(-cudc) %>%}\StringTok{ }\NormalTok{.$data %>%}\StringTok{ }\NormalTok{NROW}
\CommentTok{#> [1] 2}
\end{Highlighting}
\end{Shaded}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-unnamed-chunk-19-1.pdf}
\caption{}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-unnamed-chunk-24-1.pdf}
\caption{}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-unnamed-chunk-31-1.pdf}
\caption{}
\end{figure}

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  pt 1
\item
  pt 2
\item
  pt 3
\item
  pt 4
\end{itemize}

>>>>

Perhaps the most fundamental element in many fields of ecology is the
individual. The number of individuals of each species in a given
location forms the basis for many sub-fields of ecology and evolution.
Some research questions necessitate collecting new data, while others
can easily take advantage of existing data. In fact, some ecology fields
are built largely on existing data, e.g., macro-ecology (Brown, 1995;
Beck et al., 2012).

Data on individuals, including which species, and where they're found,
can be used for a large number of research questions. Biodiversity
records have been used for a suite of other use cases: validating
habitat suitability models with real occurrence data (Ficetola et al.,
2014); ancestral range reconstruction (Ferretti et al., 2015; Mar{í}a
Mendoza et al., 2015); development of invasive species watch lists
(Faulkner et al., 2014); evaluating risk of invasive species spread
(Febbraro et al., 2013); and effects of climate change on future
biodiversity (Brown et al., 2015).

In addition to wide utility, this data is important for conservation.
Biodiversity loss is one of the greatest challenges of our time (Pimm et
al., 2014), and some have called this the sixth great mass extinction
(Ceballos et al., 2015). Given this challenge there is a great need for
data on specimen records, whether collected from live sightings in the
field or specimens in museums.

There are many online services that collect and maintain specimen
records. However, Global Biodiversity Information Facility (hereafter,
GBIF, \url{http://www.gbif.org}) is the largest collection of
biodiversity records globally, currently with 640 million records, 1.6
million taxa, 15,000 datasets from 780 publishers (current as of
2015-11-25). Many large biodiversity warehouses such as iNaturalist
(\url{http://www.inaturalist.org}), VertNet (\url{http://vertnet.org}),
and USGS's Biodiversity Information Serving Our Nation (BISON;
\url{http://bison.usgs.ornl.gov}) all feed into GBIF.

Herein, we describe the \texttt{rgbif} software library (Chamberlain et
al.) for working with GBIF data in the R programming environment (R Core
Team, 2014). R is a widely used language in academia, and in non-profit
and private sectors. Importantly, R makes it easy to do all of the steps
of the research process, including data management, data manipulation
and cleaning, statistics, and vizualization. Thus, an R client for
getting GBIF data is a powerful tool to facilitate reproducible
research.

@@

The \texttt{rgbif} package is nearly completely written in R (a small
Javascript library is included for reading well known text), uses an
\href{http://choosealicense.com/licenses/mit/}{MIT license} to maximize
use everywhere. \texttt{rgbif} is developed publicly on GitHub at
\href{https://github.com/ropensci/rgbif}{\url{https://github.com/ropensci/rgbif}},
where development versions of the package can be installed, and bugs and
feature requests reported. Stable versions of \texttt{rgbif} can be
installed from
\href{https://cran.rstudio.com/web/packages/rgbif/}{CRAN}, the
distribution network for R packages. \texttt{rgbif} is part of the
rOpenSci project (\url{http://ropensci.org}), a developer network making
R software to facilitate reproducible research.

@@

\texttt{rgbif} is designed following the
\href{http://www.gbif.org/developer/summary}{GBIF Application
Programming Interface}, or API. The GBIF API has four major components:
registry, taxonomic names, occurrences, and maps. We also include
functions to interface with the OAI-PMH GBIF service; only dataset
information is available vis this service, however. We ignore maps in
\texttt{rgbif} as it is concerned with generating maps primarily for web
applications. \texttt{rgbif} has a suite of functions dealing with each
of registry, taxonomic names, and occurrences - we'll go through each in
turn describing design and example usage.

@@

With each request \texttt{rgbif} makes to GBIF's API, we send request
headers that tell GBIF that the request is coming from \texttt{rgbif},
including what version of the package. This helps GBIF know what
proportion of requests are coming from this package, and therefore R, as
most requests likely will come from \texttt{rgbif}; this information is
helpful for them in thinking about how people are using GBIF data.

@@

And dataset information in general is available via the OAI-PMH service,
functions in \texttt{rgbif} prefixed with \texttt{gbif\_oai\_}.

Datasets are owned by organizations. Organizations are endorsed by nodes
to share datasets with GBIF. Datasets are published through
institutions, which may be hosted at another organization. A network is
a group of datasets (managed by GBIF). Datasets are the units that
matter the most with respect to registry information, while
installations, networks, nodes, and organizations are simply higher
level organizational structure.

@@

See also \texttt{datasets()} and \texttt{dataset\_suggest()} for
searching for datasets.

@@

Dataset metrics are another useful way of figuring out what datasets you
may want to use. One drawback is that these metrics data are only
available for datasets of type \emph{checklist}, but there are quite a
lot of them (2490).

@@

One of the quickest ways to search for names is using
\texttt{name\_suggest()}, which does a very quick search and returns
minimal data. Here, we're searching for the query tem \emph{Pum}, and we
get back many names:

@@

The download API syntax is similar to the occurrence search API in that
the same parameters are used, but the way in which the query is defined
is different. For example, in the download API you can do greater than
searches (i.e., \texttt{latitude \textgreater{} 50}), whereas you can
not do that in the occurrence search API. Thus, unfortnately, we
couldn't make the query interace exactly the same for both search and
download functions.

Using the download service can consist of as few as three steps: 1)
Request data via a search; 2) Download data; 3) Import data into R.

@@

\paragraph{Search for occurrences}\label{search-for-occurrences}

@@

The output of \texttt{occ\_search()} borrows the tidy
\texttt{data.frame} idea from the \texttt{dplyr} R package, so that no
matter how large the \texttt{data.frame}, the output is easily assessed
because only a few of the records (rows) are shown, only a few columns
are shown (with others shown in name only), and metadata is shown on top
of the \texttt{data.frame} to indicate data found and returned, media
records found, unique taxonomic hierarchies returned, and the query
executed.

The output of these examples, except one, aren't shown, but all run
correctly.

Search by species name, using \texttt{name\_backbone()} first to get key

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{(key <-}\StringTok{ }\KeywordTok{name_suggest}\NormalTok{(}\DataTypeTok{q =} \StringTok{'Helianthus annuus'}\NormalTok{, }\DataTypeTok{rank =} \StringTok{'species'}\NormalTok{)$key[}\DecValTok{1}\NormalTok{])}
\CommentTok{#> [1] 3119195}
\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =} \NormalTok{key, }\DataTypeTok{limit =} \DecValTok{2}\NormalTok{)}
\CommentTok{#> Records found [21577] }
\CommentTok{#> Records returned [2] }
\CommentTok{#> No. unique hierarchies [1] }
\CommentTok{#> No. media records [2] }
\CommentTok{#> Args [taxonKey=3119195, limit=2, offset=0, fields=all] }
\CommentTok{#> First 10 rows of data}
\CommentTok{#> }
\CommentTok{#>                name        key decimalLatitude decimalLongitude}
\CommentTok{#> 1 Helianthus annuus 1143516596        35.42767        -105.0688}
\CommentTok{#> 2 Helianthus annuus 1095851641         0.00000           0.0000}
\CommentTok{#> Variables not shown: issues (chr), datasetKey (chr), publishingOrgKey}
\CommentTok{#>      (chr), publishingCountry (chr), protocol (chr), lastCrawled (chr),}
\CommentTok{#>      lastParsed (chr), extensions (chr), basisOfRecord (chr), taxonKey}
\CommentTok{#>      (int), kingdomKey (int), phylumKey (int), classKey (int), orderKey}
\CommentTok{#>      (int), familyKey (int), genusKey (int), speciesKey (int),}
\CommentTok{#>      scientificName (chr), kingdom (chr), phylum (chr), order (chr),}
\CommentTok{#>      family (chr), genus (chr), species (chr), genericName (chr),}
\CommentTok{#>      specificEpithet (chr), taxonRank (chr), dateIdentified (chr), year}
\CommentTok{#>      (int), month (int), day (int), eventDate (chr), modified (chr),}
\CommentTok{#>      lastInterpreted (chr), references (chr), identifiers (chr), facts}
\CommentTok{#>      (chr), relations (chr), geodeticDatum (chr), class (chr), countryCode}
\CommentTok{#>      (chr), country (chr), rightsHolder (chr), identifier (chr),}
\CommentTok{#>      verbatimEventDate (chr), datasetName (chr), gbifID (chr),}
\CommentTok{#>      verbatimLocality (chr), collectionCode (chr), occurrenceID (chr),}
\CommentTok{#>      taxonID (chr), recordedBy (chr), catalogNumber (chr),}
\CommentTok{#>      http...unknown.org.occurrenceDetails (chr), institutionCode (chr),}
\CommentTok{#>      rights (chr), occurrenceRemarks (chr), identificationID (chr),}
\CommentTok{#>      elevation (dbl), elevationAccuracy (dbl), stateProvince (chr),}
\CommentTok{#>      recordNumber (chr), locality (chr), municipality (chr), language}
\CommentTok{#>      (chr), type (chr), ownerInstitutionCode (chr), identifiedBy (chr)}
\end{Highlighting}
\end{Shaded}

Instead of getting a taxon key first, you can search for a name directly

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{scientificName =} \StringTok{'Ursus americanus'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Search for many species

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{splist <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{'Cyanocitta stelleri'}\NormalTok{, }\StringTok{'Junco hyemalis'}\NormalTok{, }\StringTok{'Aix sponsa'}\NormalTok{)}
\NormalTok{keys <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(splist, function(x) }\KeywordTok{name_suggest}\NormalTok{(x)$key[}\DecValTok{1}\NormalTok{], }\DataTypeTok{USE.NAMES =} \OtherTok{FALSE}\NormalTok{)}
\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =} \NormalTok{keys, }\DataTypeTok{limit =} \DecValTok{5}\NormalTok{, }\DataTypeTok{return =} \StringTok{'data'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Spatial search, based on well known text format, or a bounding box set
of four coordinates

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# well known text}
\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \StringTok{'POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))'}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{20}\NormalTok{)}
\CommentTok{# bounding box}
\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \KeywordTok{c}\NormalTok{(-}\FloatTok{125.0}\NormalTok{,}\FloatTok{38.4}\NormalTok{,-}\FloatTok{121.8}\NormalTok{,}\FloatTok{40.9}\NormalTok{), }\DataTypeTok{limit =} \DecValTok{20}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Get only occurrences with lat/long data

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{hasCoordinate =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Get only those occurrences with spatial issues

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{hasGeospatialIssue =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{out <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{issue =} \StringTok{'DEPTH_UNLIKELY'}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{500}\NormalTok{)}
\KeywordTok{NROW}\NormalTok{(out)}
\CommentTok{#> [1] 4}
\NormalTok{out %>%}\StringTok{ }\KeywordTok{occ_issues}\NormalTok{(-cudc) %>%}\StringTok{ }\NormalTok{.$data %>%}\StringTok{ }\NormalTok{NROW}
\CommentTok{#> [1] 2}
\end{Highlighting}
\end{Shaded}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-unnamed-chunk-25-1.pdf}
\caption{}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-unnamed-chunk-30-1.pdf}
\caption{}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-unnamed-chunk-37-1.pdf}
\caption{}
\end{figure}

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  pt 1
\item
  pt 2
\item
  The \texttt{rgbif} package has already been used in 22 scholarly
  publications (as of 2015-11-14).
\item
  pt 4
\end{itemize}
